---
title: 'Lession 3: Manipulating Data by Daniel Stoxreiter'
---

## Converting data from PNG to Excel
I started the homework with taking pictures from the datasets in question.
To convert the data I used OCR Engine (https://www.onlineocr.net/) and converted it to Excel. 

## Regular Expressions using Atom
I copied the content from the excel sheet to Atom and applied the following regular expressions:

```{r}
#* Step One: 
#Find: (\d+), Replace: $1 (removes the comma of Size)

#* Step Two: 
#Find: \.$ Replace: empty (removes the dot of Points for integer values)

#* Step Three: 
#Find: \s Replace: , (removes all whitespaces and replaces them with comma)

#* Step Four:
#Find: (\d+,BCE) Replace: -$1 (will add a minus at the beginning of the line for all rows with BCE)

#* Step Five:
#Find: $ Replace: ,East (will add a new obseveation "East", same for "West")
```

## Manual adjastements:
All rows with multiple city names were manually adjusted and copied into a seperate row.
I tried some code examples with tidyr but I could not achieve the desirable results.

## Script examples for data plotting

### War Making Capacity East vs. West

```{r }
library(readr)
settlementSorted <- read_csv("~/R_course/lession3/settlementSorted.csv")

#creates variables east and west to create two data sets
east <- settlementSorted[settlementSorted$Region == "East", ]
west <- settlementSorted[settlementSorted$Region == "West", ]
plot_colors <- c("blue","red")

#plots east against west via date and points
plot(x = east$Date, y = east$Points, type ="o", col = plot_colors[1], xlab="Year", ylab="War making capacity", main="War of Cities")
lines(x = west$Date, y = west$Points, type = "o", col = plot_colors[2])
legend(-4000, 250, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```


```{r}
#creates variable to seperate BCE and CE to allow better data comparison
westBCE <- west[west$`BCE-CE` == "BCE", ]
eastBCE <- east[east$`BCE-CE` == "BCE", ]

#plots east against west in date in points for BCE dates only
plot(x = eastBCE$Date, y = eastBCE$Points, type ="o", col = plot_colors[1], xlab="Year", ylab="War making capacity", main="War of Cities")
lines(x = westBCE$Date, y = westBCE$Points, type = "o", col = plot_colors[2])
legend(-4000, 4, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

```{r}
#creates variables for CE dates
westCE <- west[west$`BCE-CE` == "CE", ]
eastCE <- east[east$`BCE-CE` == "CE", ]

#plots east against west in date and points between date 0 and 2000
plot(x = eastCE$Date, y = eastCE$Points / 1000, type ="o", col = plot_colors[1], xlab="Year", ylab="War making capacity", main="War of Cities")
lines(x = westCE$Date, y = westCE$Points / 1000, type = "o", col = plot_colors[2])
legend(100, 0.25, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

```{r}
#creates variables to seperate dates between 0 and 1700 from dates between 1700 and 2000
westEarlyCE <- westCE[westCE$Date < 1700, ]
westLateCE <- westCE[westCE$Date > 1700, ]
eastEarlyCE <- eastCE[eastCE$Date < 1700, ]
eastLateCE <- eastCE[eastCE$Date > 1700, ]

#plots east against west in date and points for dates > 0 and < 1700
plot(x = eastEarlyCE$Date, y = eastEarlyCE$Points, type ="o", col = plot_colors[1], xlab="Year", ylab="War making capacity", main="War of Cities")
lines(x = westEarlyCE$Date, y = westEarlyCE$Points, type = "o", col = plot_colors[2])
legend(100, 6, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

```{r}

#plots east against west in date and Points for dates > 1700
plot(x = eastLateCE$Date, y = eastLateCE$Points, type ="o", col = plot_colors[1], xlab="Year", ylab="War making capacity", main="War of Cities")
lines(x = westLateCE$Date, y = westLateCE$Points, type = "o", col = plot_colors[2])
legend(1800, 250, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

```{r}
#plots east against west in date in Points for BCE dates only
plot(x = eastBCE$Date, y = eastBCE$Points, type ="o", col = plot_colors[1], xlab="Year", ylab="War making capacity", main="War of Cities")
lines(x = westBCE$Date, y = westBCE$Points, type = "o", col = plot_colors[2])
legend(-4000, 4, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

### Settlements of East and West

```{r}
#plots east against west in date and size for CE dates only
plot(x = eastCE$Date, y = eastCE$Size / 1000, type ="o", col = plot_colors[1], xlab="Year", ylab="Settlement Size", main="Biggest Settlements")
lines(x = westCE$Date, y = westCE$Size / 1000, type = "o", col = plot_colors[2])
legend(100, 25000, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

```{r}
#creates variables to seperate dates between 0 and 1700
westEarlyCE <- westCE[westCE$Date < 1700, ]
eastEarlyCE <- eastCE[eastCE$Date < 1700, ]

#plots east against west in date and size for dates > 0 and < 1700
plot(x = eastEarlyCE$Date, y = eastEarlyCE$Size / 1000, type ="o", col = plot_colors[1], xlab="Year", ylab="Settlement Size", main="Biggest Settlements")
lines(x = westEarlyCE$Date, y = westEarlyCE$Size / 1000, type = "o", col = plot_colors[2])
legend(100, 700, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

```{r}
#creates variables to seperate dates between 1700 and 2000
westLateCE <- westCE[westCE$Date > 1700, ]
eastLateCE <- eastCE[eastCE$Date > 1700, ]
#plots east against west in date and size for dates > 1700
plot(x = eastLateCE$Date, y = eastLateCE$Size / 1000, type ="o", col = plot_colors[1], xlab="Year", ylab="Settlement Size", main="Biggest Settlements")
lines(x = westLateCE$Date, y = westLateCE$Size / 1000, type = "o", plot_colors[2])
legend(1800, 25000, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```



```{r}
east <- settlementSorted[settlementSorted$Region == "East", ]
west <- settlementSorted[settlementSorted$Region == "West", ]

plot(x = east$Date, y = east$Size / 1000, type ="o", col = plot_colors[1], xlab="Year", ylab="Settlement Size", main="Biggest Settlements")
lines(x = west$Date, y = west$Size / 1000, type = "o", col = plot_colors[2])
legend(-4000, 25000, c("east","west"), cex=0.9, col = plot_colors, pch=21:22, lty=1:2)
```

## Trying to find the biggest settlements if the War Making Capacity of the East or West was greater
I had problems to get to the right solution. I tried various examples but the outcome is not very satisfying.


```{r}
#creates variables east and west to create two data sets
east <- settlementSorted[settlementSorted$Region == "East", ]
west <- settlementSorted[settlementSorted$Region == "West", ]

eastGreatSettlement <- settlementSorted$Settlement[east$Points > west$Points & east$Size > west$Size]

westGreatSettlement <- settlementSorted$Settlement[east$Points < west$Points & east$Size < west$Size]

eastGreatSettlement

westGreatSettlement

```

```{r}
#eastGreatSettlement <- east$Settlement[(east$Points > west$Points) & (east$Size > west$Size)]
```

```{r}
#eastGreatSettlement <- east$Points > west$Points & east$Size > west$Size
#westGreatSettlement <- east$Points < west$Points & east$Size < west$Size

#eastGreatSettlement
```

