---
title: "Visualizing Geographical Data"
author: "Daniel Stoxreiter"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

**NB:** The worksheet has beed developed and prepared by Maxim Romanov for the course "R for Historical Research" (U Vienna, Spring 2019)

## Necessarey libraries

We will need some new libraries that are developed for working with maps. You may want to install those libraries outside of the notebook, so that it does not try to reinstall them everytime you `knit` it.

```{r}
#install.packages(c("cowplot", "googleway", "ggplot2", "ggrepel", "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata", "rnaturalearthhires", "rgeos"))
```


```{r}
library(tinytex)
library("ggplot2")
theme_set(theme_bw())
library("sf")

```

The libraries below have the necessary geographical data, so let's load them and create a working variable.

```{r}
library("rnaturalearth")
library("rnaturalearthdata")

#world <- ne_countries(scale = "large", returnclass = "sf")
world <- ne_countries(scale = "medium", returnclass = "sf")
```

Now we can create our first and simple map with just a couple of lines of code. See how easy it is.

```{r}
ggplot(data = world) +
    geom_sf()

```

We can add additional information to our map to make it more informative. Look at the `ggtitle()`: we can combine necessary numbers with textual annotations.

```{r}
ggplot(data = world) +
    geom_sf() +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("World map", subtitle = paste0("(", length(unique(world$name)), " countries)"))

```

`Color` and `fill` determine the colors of "borders" and "areas" within. 

```{r}
ggplot(data = world) + 
    geom_sf(color = "grey", fill = "grey")

```

Since we have some demographic data, we can visualize it with colors. You can `google` other color scales.

```{r}
ggplot(data = world) +
    geom_sf(aes(fill = pop_est)) +
    scale_fill_viridis_c(option = "magma", trans = "sqrt")

```

We can also select a projection like shown in the code below. For more details, check references given below.

```{r}
ggplot(data = world) +
    geom_sf(color = "grey", fill = "grey") +
    coord_sf(crs = st_crs(3035))
```

We can focus our map on a specific area of the globe.

```{r}

xlim=c(85,130); ylim=c(10,45)

ggplot(data = world) +
    geom_sf(color="grey", fill="grey") +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE)

```

And add a scale and a North arrow, to make it look a bit cooler.

```{r message=FALSE}
library("ggspatial")

xlim=c(-12,40); ylim=c(30,60)
ggplot(data = world) +
    geom_sf(fill="grey60", color="grey60") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5), panel.background = element_rect(fill = "grey90"))

```

## Data

Let's get some data first. The data comes from: <http://atlantides.org/downloads/pleiades/dumps/>. Can you find the description of data? (If description is already available it may make it easier for you to figure out what can be done with it.)

```{r message=TRUE}

data_link <- "http://atlantides.org/downloads/pleiades/dumps/pleiades-locations-latest.csv.gz"
download.file(data_link,"pleiades-locations-latest.csv.gz")
locsRaw <- read.csv(gzfile("pleiades-locations-latest.csv.gz"),stringsAsFactors=F,header=T,sep=',')

```

The data contains *geolocations*, which are also categorized chronologically, The chronology is rather tricky: there are several main periods, but they are not used discretionally as many places often exist in multiple periods. These look as follows:

* `archaic` (750-550BC)
* `classical` (550-330BC)
* `hellenistic-republican` (330-30BC)
* `roman` (30BC-300CE)
* `late-antique` (300-640CE)

You can also try to split this data into other periods based on the `minDate` and `maxDate`, but that is a bit more complicated.

Let's try to add it to our base map.

```{r message=FALSE, warning=FALSE}

xlim <- c(-12,55); ylim=c(20,60)

plot <- ggplot(data = world) +
    geom_sf(fill="grey70", color="grey70") +
    geom_point(data=locsRaw,color="red",alpha=.1,size=.5,aes(y=reprLat,x=reprLong)) +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("Data from Pleiades", subtitle = "Density of Geographical Information")

plot
```

Let's check what is in the data (always try to find the official description of data!).

```{r}
summary(locsRaw)
```

We check what types of things are mappable. In this dataset, types of objects are described in the `featureType` column.

```{r}
unique(locsRaw$featureType)
```

You can use this line of code to check what kind of information is stored in other columns.

Now we can select some specific feature we are interested in and add this data to our cartogram. As you can see, however, the data is quite messy and makes things rather difficult. So, for filtering our data it is best to use `regular expressions` like shown below. You can use `nrow()` to count how many of those items are in the data mapping some low frequency objects may not be very informative.

```{r}
locs=locsRaw[ with(locsRaw, grepl("\\bfort\\b",featureType)),]
nrow(locs)
```

We now can add anther layer of some specific data. Let's try *forts*, which we have just created above.

```{r message=FALSE, warning=FALSE}

xlim=c(-12,55); ylim=c(20,60)

ggplot(data = world) +
    geom_sf(fill="grey70", color="grey70") +
    geom_point(data=locsRaw,color="grey20",alpha=.1,size=.1,aes(y=reprLat,x=reprLong)) +
    geom_point(data=locs,color="red",alpha=.75,size=.5,aes(y=reprLat,x=reprLong)) +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("") + ylab("") +
    ggtitle("Forts from Pleiades", subtitle = "Density of Geographical Information")

```

Let's check whether there are/were any forts on the territory of modern-day Austria.

```{r warning=FALSE}
#looks for austria in the world dataset
austria <- subset(world, admin == "Austria")
# defines the dataset
ourPlot <- ggplot(data = austria) +
  # sets the color of the map in different grey shades for borders and fill
    geom_sf(fill="grey70", color="grey50") +
  #adds geom point from data locsRaw and log lat from this file as variables
    geom_point(data=locsRaw,color="grey20",alpha=.25,size=.15,aes(y=reprLat,x=reprLong)) +
  #the earlier subset "locs" provides data for forts only
    geom_point(data=locs,color="red",alpha=.75,size=1,aes(y=reprLat,x=reprLong)) +
  #adds another geom point for showing vienna on the map
    geom_point(color="black",alpha=1,shape=1, size=3,aes(y=48.210033,x=16.363449)) +
  # adds text and labels were Vienna text is slightly outside the coordinates
    geom_text(x=16.363449+.5, y=48.210033+.1, label="vienna", color="black", size=4) +
    coord_sf(xlim = c(7.5,17.5), ylim = c(46,50), expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "bl", width_hint = 0.25) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Ancient and Early Medieval Forts", subtitle = "Zooming on Modern Austria")

ourPlot
# saves the graph in a file
ggsave(file="forts_in_austria.png",plot=ourPlot,dpi=300,width=7,height=5)

```

By the way, we can store the plot into a variable and then reuse it. This is convenient for saving: we can use `ggsave()` as shown above to save our plot into a separate image. Check the folder with your R Notebook there should be a file called `forts_in_austria.png`. There is a variety of parameters that you can adjust to get the exact quality of your image that you need. Check `?ggsave` for more details (or `google` it).

Now, let's do some exercises.

(@) Can you check which historical period those forts belong to? Do not forget to change the legend of your map accordingly. (*Hint:* check the note on historical periods above).

* `archaic` (750-550BC)
* `classical` (550-330BC)
* `hellenistic-republican` (330-30BC)
* `roman` (30BC-300CE)
* `late-antique` (300-640CE)

```{r}

austria <- subset(world, admin == "Austria")

locs=locsRaw[ with(locsRaw, grepl("\\bfort\\b",featureType)),]
locsPer=locs[ with(locs, grepl("\\broman\\b",timePeriodsKeys)),]

ourPlot <- ggplot(data = austria) +
    geom_sf(fill="grey70", color="grey50") +
    geom_point(data=locsRaw,color="grey20",alpha=.25,size=.15,aes(y=reprLat,x=reprLong)) +
    #geom_point(data=locs,color="red",alpha=.75,size=1,aes(y=reprLat,x=reprLong)) +
    geom_point(data=locsPer,color="blue",alpha=.75,size=1,aes(y=reprLat,x=reprLong)) +
    geom_point(color="black",alpha=1,shape=1, size=3,aes(y=48.210033,x=16.363449)) +
    geom_text(x=16.363449+.5, y=48.210033+.1, label="vienna", color="black", size=4) +
    coord_sf(xlim = c(7.5,17.5), ylim = c(46,50), expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "bl", width_hint = 0.25) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Ancient and Early Medieval Forts", subtitle = "Zooming on Modern Austria")

ourPlot
ggsave(file="forts_in_austria_period_period.png",plot=ourPlot,dpi=300,width=7,height=5)

```

(@) Can you create the map of (Hadrian's Wall)[https://en.wikipedia.org/wiki/Hadrian's_Wall]? Try to make the map as informative as possible (you can use a map from Wikipedia as a model).

```{r}
# narrow the data down to long lat were the wall is
library(dplyr)
locsRaw
hadrian <- locsRaw %>%
  filter(reprLong > -5 & reprLong < 1) %>%
  filter(reprLat > 54.5 & reprLat < 55.5)
```


```{r}
library(ggplot2)
library(ggspatial)
gb <- subset(world, admin == "United Kingdom")

locs=locsRaw[ with(locsRaw, grepl("\\bfort\\b",featureType)),]
locsWall=locsRaw[ with(locsRaw, grepl("\\bunknown\\b",featureType)),]

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey70", color="grey50") +
    #geom_point(data=locsRaw,color="pink",alpha=.25,size=.15,aes(y=reprLat,x=reprLong)) +
    geom_point(data=locs,color="red",alpha=.75,size=1,aes(y=reprLat,x=reprLong)) +
    geom_point(data=locsWall,color="blue",alpha=.75,size=1,aes(y=reprLat,x=reprLong)) +
    #geom_point(color="black",alpha=1,shape=1, size=3,aes(y=48.210033,x=16.363449)) +
    #geom_text(x=16.363449+.5, y=48.210033+.1, label="vienna", color="black", size=4) +
    coord_sf(xlim = c(-4,0), ylim = c(53,57), expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "bl", width_hint = 0.25) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Hadrians Wall or THE WALL", subtitle = "watch out for white walkers")

ourPlot
ggsave(file="forts_in_hadrians_wall.png",plot=ourPlot,dpi=300,width=7,height=5)

```

(@) Explore Pleiades dataset and create three maps of something interesting that you discover.

(@) Plot 1: *Plot that shows Europe, North Africa and parts of Asia with an timelaps of forts that were build in different timeperiods.* (but also do not forget to add all necessary information into the map/cartogram).

```{r}
#install.packages("gifski")
library(gifski)
library(gganimate)
library(ggplot2)
library(ggspatial)
library(dplyr)

# This first part search for regular expression for all time periods of "timePeriodsKeys" that are relevant.

locsPeriods=locsRaw[ with(locsRaw, grepl("\\broman\\b|\\bclassical\\b|\\barchaic\\b|\\blate-antique\\b|\\hellenistic-republican\\b",timePeriodsKeys)),]

# The second part searches for a specifc first letter in "timePeriods" or a specifc first letter plus any other following letter. E.g. R = Roman
# RL = Roman,Late-Antique

locsArchaic=locsPeriods[ with(locsPeriods, grepl("^A|A\\w+",timePeriods)),]
locsRoman=locsPeriods[ with(locsPeriods, grepl("^R|R\\w+",timePeriods)),]
locsClassical=locsPeriods[ with(locsPeriods, grepl("^C|C\\w+",timePeriods)),]
locsHellenisticRepublican=locsPeriods[ with(locsPeriods, grepl("^H|H\\w+",timePeriods)),]
locsLateAntique=locsPeriods[ with(locsPeriods, grepl("^L|L\\w+",timePeriods)),]

# Before I could work on the plot I had to clean the data and add a new column called "epoch". The new variable assigns observatoins to one specific timeperiod.
# I assumed that the varialbe "timePeriods" does provide shortcuts for timePeriodsKeys and the first letter in "timePeriods" would be the period in which a fort, # or settlement, or any sight was built first. E.g. "RL" = "Romen,Late-Antique" will be assiged to the value "roman". 

locsArchaic <- locsArchaic %>%
  select(uid) %>%
  mutate(epoch = "archaic")

locsRoman <- locsRoman %>%
  select(uid) %>%
  mutate(epoch = "roman")

locsClassical <- locsClassical %>%
  select(uid) %>%
  mutate(epoch = "classical")

locsHellenisticRepublican <- locsHellenisticRepublican %>%
  select(uid) %>%
  mutate(epoch = "hellenistic-republican")

locsLateAntique <- locsLateAntique %>%
  select(uid) %>%
  mutate(epoch = "late-antique")

# The following section joins the resulting different data frames fully into one

joinLocs <- full_join(locsClassical, locsArchaic)
joinLocs <- full_join(joinLocs, locsRoman)
joinLocs <- full_join(joinLocs, locsLateAntique)
joinLocs <- full_join(joinLocs, locsHellenisticRepublican)

# The next section left_joins the larger database in "locsPeriods" with the newly "joinLocs" data frame.

joinedLocs <- left_join(locsPeriods, joinLocs, by = "uid")

# Finally filter for specifc timeperiods to avoid NAs

locsNicelyCleaned <- joinedLocs %>%
  filter(epoch == c("classical", "roman", "late-antique", "hellenistic-republican", "archaic"))
```


```{r}
# searches for "fort" with regular expression to receive forts only
locs=locsNicelyCleaned[ with(locsNicelyCleaned, grepl("\\bfort\\b",featureType)),]

# Plotting

# creates a plot with a world map
ourPlot <- ggplot(data = world) +
    # creates the colors for the world map
    geom_sf(fill="grey70", color="grey50") +
    # creates plottings points from the "locs" dataframe and adds a dot on specific coordinates
    geom_point(data=locs,alpha=.7,size=2,aes(y=reprLat,x=reprLong,color=epoch)) +
    # defines the area of the world map that is displayed
    coord_sf(xlim = c(-15,50), ylim = c(20,75), expand = FALSE) +
    # defines the theme of the plot
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    # assigns names to x and y axis
    xlab("Longitude") + ylab("Latitude") +
    # creates a annotation scale
    annotation_scale(location = "bl", width_hint = 0.25) +
    # creates an north arrow for the annotation
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    # creates the title for the plot
    ggtitle("Forts timelaps {closest_state}", subtitle = "Frame {frame} of {nframes}")+
    # changes the title for the colors which are the different timeperiods of the epoch column
    labs(color = "Timeperiods")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/  
    # animation that uses the  new "epoch" column to display forts of various timperiods one by one
    # transition_states(epoch,
    #                 transition_length = 3,
    #                 state_length = 100,
    #                 wrap = FALSE)+
    # enter_fade()+
    # exit_fade()

ourPlot
ggsave(file="forts_in_different_periods.png",plot=ourPlot,dpi=300,width=7,height=5)
```


(@) Plot 2: *This graphs shows Churches that were build in different timeperiods. (The animation/transition is commented out due to long rendering times)* (but also do not forget to add all necessary information into the map/cartogram).

```{r}
locsChurch=locsNicelyCleaned[ with(locsNicelyCleaned, grepl("\\bchurch\\b",featureType)),]

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey70", color="grey50") +
    geom_point(data=locsChurch,size=2,aes(y=reprLat,x=reprLong,color=epoch)) +
    coord_sf(xlim = c(-15,50), ylim = c(20,75), expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "bl", width_hint = 0.25) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Church timelaps {closest_state}", subtitle = "Frame {frame} of {nframes}")+
    labs(color = "Timeperiods")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/  
    # transition_states(epoch,
    #                 transition_length = 3,
    #                 state_length = 100,
    #                 wrap = FALSE)+
    # enter_fade()+
    # exit_fade()

ourPlot
ggsave(file="churches_in_different_timeperiods.png",plot=ourPlot,dpi=300,width=7,height=5)
```

(@) Plot 3: *This plot shows Churches in red dots and Settlements (The animation/transition is commented out due to long rendering times)* (but also do not forget to add all necessary information into the map/cartogram).

```{r}
locsSettlement=locsNicelyCleaned[ with(locsNicelyCleaned, grepl("\\bsettlement\\b",featureType)),]
locsChurch=locsNicelyCleaned[ with(locsNicelyCleaned, grepl("\\bchurch\\b",featureType)),]

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey70", color="grey50") +
    geom_point(data=locsChurch,color="red",size=5,aes(y=reprLat,x=reprLong)) +
    geom_point(data=locsSettlement,alpha=.7,size=2,aes(y=reprLat,x=reprLong,color=epoch)) +
    coord_sf(xlim = c(-15,50), ylim = c(20,75), expand = FALSE) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "bl", width_hint = 0.25) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Settlements and Churches timelaps of {closest_state} period", subtitle = "Frame {frame} of {nframes}")+
    labs(color = "Timeperiods")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
    # transition_states(epoch,
    #                 transition_length = 3,
    #                 state_length = 100,
    #                 wrap = FALSE)+
    # enter_fade()+
    # exit_fade()

ourPlot
ggsave(file="settlements_in_different_timeperiods.png",plot=ourPlot,dpi=300,width=7,height=5)
```


## `Europop` package

The `europop` package has information on population of european cities, which we can now map. There are two main tables:  `europop` with the historical populations of European cities, and `city_coords`âwith the latitudes and longitudes of those cities. With this new data you now should be able to map European cities. Your task is as follows:

* create maps of 5 different periods (i.e., 5 maps)
* colorcode your map by regions
* add the names of the top 10 largest cities to your maps
* add relevant legends to your maps
* you may need to adjust your legend (`google` it!) 

```{r}
#install.packages("europop")
library(dplyr)
library(tidyverse)
library(europop)
library(ggplot2)
library(ggspatial)
library(gganimate)
biggestCities <- europop %>%
  filter(population >= 305 & year == 1800)

new_data <- data.frame("region" = c("Scandinavia", "England and Wales", "Scotland", "Ireland", "The Netherlands", "Belgium", "Germany", "France", "Switzerland", "Northern Italy", "Central Italy", "Souther Italy", "Spain", "Portual", "Austria and Czechoslovakia", "Poland"), "long" = c(15.150000, -1.210000, -3.900000, -7.480000, 6.220000, 4.810000, 9.800000, 2.250000, 8.060000, 10.170000, 12.354946, 16.581014, -2.939996, -8.346877, 16.456718, 17.009529), "lat" = c(61.400000, 51.980000, 56.960000, 52.690000, 52.990000, 50.900000, 50.500000, 48.290000, 46.890000, 45.337603, 42.139779, 39.170254, 39.664747, 48.686041, 48.509328, 51.862042))

europopCoor <- left_join(europop, new_data, by = "region")
```


```{r}
ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey80", color="grey50") +
    geom_count(data=europopCoor,alpha=.6,aes(y=lat,x=long,color=region,size=population), na.rm=TRUE) +
    geom_count(data=europopCoor,color="red",alpha=.6,aes(y=63,x=0,size=population), na.rm=TRUE) +  
    coord_sf(xlim = c(-13,30), ylim = c(35,70), expand = FALSE) +
    geom_text(x=-0.241680+.5, y=51.528558+.1, label="London", color="black", size=4) +
    geom_text(x=2.2770205+.5, y=48.858837+.1, label="Paris", color="black", size=4) +
    geom_text(x=14.1765627+.5, y=40.853912+.1, label="Napoli", color="black", size=4) +
    geom_text(x=4.828583+.5, y=52.354627+.1, label="Amsterdam", color="black", size=4) +
    geom_text(x=12.395915+.5, y=41.909986+.1, label="Rome", color="black", size=4) +
    geom_text(x=-6.385785+.5, y=53.324238+.1, label="Dublin", color="black", size=4) +
    geom_text(x=-3.819619+.5, y=40.437869+.1, label="Madrid", color="black", size=4) +
    geom_text(x=16.363449+.5, y=48.210033+.1, label="vienna", color="black", size=4) +
    geom_text(x=13.284650+.5, y=52.506761+.1, label="Berlin", color="black", size=4) +
    geom_text(x=-9.195222+.5, y=38.743621+.1, label="Lisboa", color="black", size=4) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Population timelaps of {closest_state}", subtitle = "Frame {frame} of {nframes}")+
    labs(color = "Regions", size = "Aggregated Population")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
    # transition_states(year,
    #                  transition_length = 5,
    #                  state_length = 100,
    #                  wrap = FALSE)+
    # enter_fade()+
    # exit_fade()

ourPlot
ggsave(file="european_population_timelaps.png",plot=ourPlot,dpi=300,width=7,height=5)

```

```{r}
library(dplyr)
library(tidyverse)
library(europop)
library(ggplot2)
library(ggspatial)
library(gganimate)
unique(europop$year)
biggestCities <- europop %>%
  filter(population >= 150)

europopRen <- europop %>%
  filter(year == 1500) %>%
  mutate(period = "Rennaissance")
  
europopRef <- europop %>%
  filter(year == 1550) %>%
  mutate(period = "Reformation")

europopWar <- europop %>%
  filter(year == 1600) %>%
  mutate(period = "30 Years War")

europopWar2 <- europop %>%
  filter(year == 1650) %>%
  mutate(period = "30 Years War")

europopRom <- europop %>%
  filter(year == 1700) %>%
  mutate(period = "Romatic")

europopRom2 <- europop %>%
  filter(year == 1750) %>%
  mutate(period = "Romatic")

europopHist <- europop %>%
  filter(year == 1800) %>%
  mutate(period = "Historicism")

europopPeriods <- full_join(europopRen, europopRef)
europopPeriods <- full_join(europopPeriods, europopWar)
europopPeriods <- full_join(europopPeriods, europopWar2)
europopPeriods <- full_join(europopPeriods, europopRom)
europopPeriods <- full_join(europopPeriods, europopRom2)
europopPeriods <- full_join(europopPeriods, europopHist)
europopPeriodsCoor <- left_join(europopCoor, europopPeriods)
```

```{r}
ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey80", color="grey50") +
    geom_count(data=europopPeriodsCoor,alpha=.6,aes(y=lat,x=long,color=region,size=population), na.rm=TRUE) +
    coord_sf(xlim = c(-13,30), ylim = c(35,70), expand = FALSE) +
    geom_text(x=-0.241680+.5, y=51.528558-.3, label="London", color="black", size=4) +
    geom_point(x=-0.241680+.5, y=51.528558, color="black", size=2, shape=2) +
    geom_text(x=2.2770205+.5, y=48.858837-.3, label="Paris", color="black", size=4) +
    geom_point(x=2.2770205+.5, y=48.858837, color="black", size=2, shape=2) +  
    geom_text(x=14.1765627+.5, y=40.853912-.3, label="Napoli", color="black", size=4) +
    geom_point(x=14.1765627+.5, y=40.853912, color="black", size=2, shape=2) +  
    geom_text(x=4.828583+.5, y=52.354627-.3, label="Amsterdam", color="black", size=4) +
    geom_point(x=4.828583+.5, y=52.354627, color="black", size=2, shape=2) +  
    geom_text(x=12.395915+.5, y=41.909986-.3, label="Rome", color="black", size=4) +
    geom_point(x=12.395915+.5, y=41.909986, color="black", size=2, shape=2) +  
    geom_text(x=-6.385785+.5, y=53.324238-.3, label="Dublin", color="black", size=4) +
    geom_point(x=-6.385785+.5, y=53.324238, color="black", size=2, shape=2) + 
    geom_text(x=-3.819619+.5, y=40.437869+.3, label="Madrid", color="black", size=4) +
    geom_point(x=-3.819619+.5, y=40.437869, color="black", size=2, shape=2) +
    geom_text(x=16.363449+.5, y=48.210033-.3, label="vienna", color="black", size=4) +
    geom_point(x=16.363449+.5, y=48.210033, color="black", size=2, shape=2) +  
    geom_text(x=13.284650+.5, y=52.506761-.3, label="Berlin", color="black", size=4) +
    geom_point(x=13.284650+.5, y=52.506761, color="black", size=2, shape=2) +  
    geom_text(x=-9.195222+.5, y=38.743621-.3, label="Lisboa", color="black", size=4) +
    geom_point(x=-9.195222+.5, y=38.743621, color="black", size=2, shape=2) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Population timelaps during {closest_state}", subtitle = "Frame {frame} of {nframes}")+
    labs(color = "Periods", size = "Aggregated Population")+
      theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
    # transition_states(period,
    #                 transition_length = 5,
    #                 state_length = 100,
    #                 wrap = FALSE)+
    # enter_fade()+
    # exit_fade()

ourPlot
ggsave(file="european_population_timelaps_periods.png",plot=ourPlot,dpi=300,width=7,height=5)
```

```{r}
Rennaissance <- europopPeriodsCoor %>%
  filter(period=="Rennaissance")

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey80", color="grey50") +
    geom_count(data=Rennaissance,alpha=.2,aes(y=lat,x=long,color=region,size=population), na.rm=TRUE) +
    coord_sf(xlim = c(-13,30), ylim = c(35,70), expand = FALSE) +
    geom_text(x=-0.241680+.5, y=51.528558-.3, label="London", color="black", size=3) +
    geom_point(x=-0.241680+.5, y=51.528558, color="black", size=1, shape=2) +
    geom_text(x=2.2770205+.5, y=48.858837-.3, label="Paris", color="black", size=3) +
    geom_point(x=2.2770205+.5, y=48.858837, color="black", size=1, shape=2) +  
    geom_text(x=14.1765627+.5, y=40.853912-.3, label="Napoli", color="black", size=3) +
    geom_point(x=14.1765627+.5, y=40.853912, color="black", size=1, shape=2) +  
    geom_text(x=4.828583+.5, y=52.354627-.3, label="Amsterdam", color="black", size=3) +
    geom_point(x=4.828583+.5, y=52.354627, color="black", size=1, shape=2) +  
    geom_text(x=12.395915+.5, y=41.909986-.3, label="Rome", color="black", size=3) +
    geom_point(x=12.395915+.5, y=41.909986, color="black", size=1, shape=2) +  
    geom_text(x=-6.385785+.5, y=53.324238-.3, label="Dublin", color="black", size=3) +
    geom_point(x=-6.385785+.5, y=53.324238, color="black", size=1, shape=2) + 
    geom_text(x=-3.819619+.5, y=40.437869+.3, label="Madrid", color="black", size=3) +
    geom_point(x=-3.819619+.5, y=40.437869, color="black", size=1, shape=2) +
    geom_text(x=16.363449+.5, y=48.210033-.3, label="vienna", color="black", size=3) +
    geom_point(x=16.363449+.5, y=48.210033, color="black", size=1, shape=2) +  
    geom_text(x=13.284650+.5, y=52.506761-.3, label="Berlin", color="black", size=3) +
    geom_point(x=13.284650+.5, y=52.506761, color="black", size=1, shape=2) +  
    geom_text(x=-9.195222+.5, y=38.743621-.3, label="Lisboa", color="black", size=3) +
    geom_point(x=-9.195222+.5, y=38.743621, color="black", size=1, shape=2) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Population during Rennaissance around 1500", subtitle = "")+
    labs(color = "Regions", size = "Aggregated Population in 1000")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
    

ourPlot
ggsave(file="european_population_period_rennaissance.png",plot=ourPlot,dpi=300,width=7,height=5)
```

```{r}
Reformation <- europopPeriodsCoor %>%
  filter(period=="Reformation")

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey80", color="grey50") +
    geom_count(data=Reformation,alpha=.2,aes(y=lat,x=long,color=region,size=population), na.rm=TRUE) +
    coord_sf(xlim = c(-13,30), ylim = c(35,70), expand = FALSE) +
    geom_text(x=-0.241680+.5, y=51.528558-.3, label="London", color="black", size=3) +
    geom_point(x=-0.241680+.5, y=51.528558, color="black", size=1, shape=2) +
    geom_text(x=2.2770205+.5, y=48.858837-.3, label="Paris", color="black", size=3) +
    geom_point(x=2.2770205+.5, y=48.858837, color="black", size=1, shape=2) +  
    geom_text(x=14.1765627+.5, y=40.853912-.3, label="Napoli", color="black", size=3) +
    geom_point(x=14.1765627+.5, y=40.853912, color="black", size=1, shape=2) +  
    geom_text(x=4.828583+.5, y=52.354627-.3, label="Amsterdam", color="black", size=3) +
    geom_point(x=4.828583+.5, y=52.354627, color="black", size=1, shape=2) +  
    geom_text(x=12.395915+.5, y=41.909986-.3, label="Rome", color="black", size=3) +
    geom_point(x=12.395915+.5, y=41.909986, color="black", size=1, shape=2) +  
    geom_text(x=-6.385785+.5, y=53.324238-.3, label="Dublin", color="black", size=3) +
    geom_point(x=-6.385785+.5, y=53.324238, color="black", size=1, shape=2) + 
    geom_text(x=-3.819619+.5, y=40.437869+.3, label="Madrid", color="black", size=3) +
    geom_point(x=-3.819619+.5, y=40.437869, color="black", size=1, shape=2) +
    geom_text(x=16.363449+.5, y=48.210033-.3, label="vienna", color="black", size=3) +
    geom_point(x=16.363449+.5, y=48.210033, color="black", size=1, shape=2) +  
    geom_text(x=13.284650+.5, y=52.506761-.3, label="Berlin", color="black", size=3) +
    geom_point(x=13.284650+.5, y=52.506761, color="black", size=1, shape=2) +  
    geom_text(x=-9.195222+.5, y=38.743621-.3, label="Lisboa", color="black", size=3) +
    geom_point(x=-9.195222+.5, y=38.743621, color="black", size=1, shape=2) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Population during Reformation around 1550", subtitle = "")+
    labs(color = "Regions", size = "Aggregated Population in 1000")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/


ourPlot
ggsave(file="european_population_period_reformation.png",plot=ourPlot,dpi=300,width=7,height=5)
```

```{r}
war <- europopPeriodsCoor %>%
  filter(period=="30 Years War")

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey80", color="grey50") +
    geom_count(data=war,alpha=.2,aes(y=lat,x=long,color=region,size=population), na.rm=TRUE) +
    coord_sf(xlim = c(-13,30), ylim = c(35,70), expand = FALSE) +
    geom_text(x=-0.241680+.5, y=51.528558-.3, label="London", color="black", size=3) +
    geom_point(x=-0.241680+.5, y=51.528558, color="black", size=1, shape=2) +
    geom_text(x=2.2770205+.5, y=48.858837-.3, label="Paris", color="black", size=3) +
    geom_point(x=2.2770205+.5, y=48.858837, color="black", size=1, shape=2) +  
    geom_text(x=14.1765627+.5, y=40.853912-.3, label="Napoli", color="black", size=3) +
    geom_point(x=14.1765627+.5, y=40.853912, color="black", size=1, shape=2) +  
    geom_text(x=4.828583+.5, y=52.354627-.3, label="Amsterdam", color="black", size=3) +
    geom_point(x=4.828583+.5, y=52.354627, color="black", size=1, shape=2) +  
    geom_text(x=12.395915+.5, y=41.909986-.3, label="Rome", color="black", size=3) +
    geom_point(x=12.395915+.5, y=41.909986, color="black", size=1, shape=2) +  
    geom_text(x=-6.385785+.5, y=53.324238-.3, label="Dublin", color="black", size=3) +
    geom_point(x=-6.385785+.5, y=53.324238, color="black", size=1, shape=2) + 
    geom_text(x=-3.819619+.5, y=40.437869+.3, label="Madrid", color="black", size=3) +
    geom_point(x=-3.819619+.5, y=40.437869, color="black", size=1, shape=2) +
    geom_text(x=16.363449+.5, y=48.210033-.3, label="vienna", color="black", size=3) +
    geom_point(x=16.363449+.5, y=48.210033, color="black", size=1, shape=2) +  
    geom_text(x=13.284650+.5, y=52.506761-.3, label="Berlin", color="black", size=3) +
    geom_point(x=13.284650+.5, y=52.506761, color="black", size=1, shape=2) +  
    geom_text(x=-9.195222+.5, y=38.743621-.3, label="Lisboa", color="black", size=3) +
    geom_point(x=-9.195222+.5, y=38.743621, color="black", size=1, shape=2) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Population during 30 Years of War from 1600 to 1650", subtitle = "")+
    labs(color = "Regions", size = "Aggregated Population in 1000")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/


ourPlot
ggsave(file="european_population_period_war.png",plot=ourPlot,dpi=300,width=7,height=5)
```


```{r}
historicism <- europopPeriodsCoor %>%
  filter(period=="Historicism")

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey80", color="grey50") +
    geom_count(data=historicism,alpha=.2,aes(y=lat,x=long,color=region,size=population), na.rm=TRUE) +
    coord_sf(xlim = c(-13,30), ylim = c(35,70), expand = FALSE) +
    geom_text(x=-0.241680+.5, y=51.528558-.3, label="London", color="black", size=3) +
    geom_point(x=-0.241680+.5, y=51.528558, color="black", size=1, shape=2) +
    geom_text(x=2.2770205+.5, y=48.858837-.3, label="Paris", color="black", size=3) +
    geom_point(x=2.2770205+.5, y=48.858837, color="black", size=1, shape=2) +  
    geom_text(x=14.1765627+.5, y=40.853912-.3, label="Napoli", color="black", size=3) +
    geom_point(x=14.1765627+.5, y=40.853912, color="black", size=1, shape=2) +  
    geom_text(x=4.828583+.5, y=52.354627-.3, label="Amsterdam", color="black", size=3) +
    geom_point(x=4.828583+.5, y=52.354627, color="black", size=1, shape=2) +  
    geom_text(x=12.395915+.5, y=41.909986-.3, label="Rome", color="black", size=3) +
    geom_point(x=12.395915+.5, y=41.909986, color="black", size=1, shape=2) +  
    geom_text(x=-6.385785+.5, y=53.324238-.3, label="Dublin", color="black", size=3) +
    geom_point(x=-6.385785+.5, y=53.324238, color="black", size=1, shape=2) + 
    geom_text(x=-3.819619+.5, y=40.437869+.3, label="Madrid", color="black", size=3) +
    geom_point(x=-3.819619+.5, y=40.437869, color="black", size=1, shape=2) +
    geom_text(x=16.363449+.5, y=48.210033-.3, label="vienna", color="black", size=3) +
    geom_point(x=16.363449+.5, y=48.210033, color="black", size=1, shape=2) +  
    geom_text(x=13.284650+.5, y=52.506761-.3, label="Berlin", color="black", size=3) +
    geom_point(x=13.284650+.5, y=52.506761, color="black", size=1, shape=2) +  
    geom_text(x=-9.195222+.5, y=38.743621-.3, label="Lisboa", color="black", size=3) +
    geom_point(x=-9.195222+.5, y=38.743621, color="black", size=1, shape=2) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Population during Historicism around 1800", subtitle = "")+
    labs(color = "Regions", size = "Aggregated Population in 1000")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/


ourPlot
ggsave(file="european_population_period_historicism.png",plot=ourPlot,dpi=300,width=7,height=5)
```

```{r}
romantic <- europopPeriodsCoor %>%
  filter(period=="Romantic")

ourPlot <- ggplot(data = world) +
    geom_sf(fill="grey80", color="grey50") +
    geom_count(data=romantic,alpha=.2,aes(y=lat,x=long,color=region,size=population), na.rm=TRUE) +
    coord_sf(xlim = c(-13,30), ylim = c(35,70), expand = FALSE) +
    geom_text(x=-0.241680+.5, y=51.528558-.3, label="London", color="black", size=3) +
    geom_point(x=-0.241680+.5, y=51.528558, color="black", size=1, shape=2) +
    geom_text(x=2.2770205+.5, y=48.858837-.3, label="Paris", color="black", size=3) +
    geom_point(x=2.2770205+.5, y=48.858837, color="black", size=1, shape=2) +  
    geom_text(x=14.1765627+.5, y=40.853912-.3, label="Napoli", color="black", size=3) +
    geom_point(x=14.1765627+.5, y=40.853912, color="black", size=1, shape=2) +  
    geom_text(x=4.828583+.5, y=52.354627-.3, label="Amsterdam", color="black", size=3) +
    geom_point(x=4.828583+.5, y=52.354627, color="black", size=1, shape=2) +  
    geom_text(x=12.395915+.5, y=41.909986-.3, label="Rome", color="black", size=3) +
    geom_point(x=12.395915+.5, y=41.909986, color="black", size=1, shape=2) +  
    geom_text(x=-6.385785+.5, y=53.324238-.3, label="Dublin", color="black", size=3) +
    geom_point(x=-6.385785+.5, y=53.324238, color="black", size=1, shape=2) + 
    geom_text(x=-3.819619+.5, y=40.437869+.3, label="Madrid", color="black", size=3) +
    geom_point(x=-3.819619+.5, y=40.437869, color="black", size=1, shape=2) +
    geom_text(x=16.363449+.5, y=48.210033-.3, label="vienna", color="black", size=3) +
    geom_point(x=16.363449+.5, y=48.210033, color="black", size=1, shape=2) +  
    geom_text(x=13.284650+.5, y=52.506761-.3, label="Berlin", color="black", size=3) +
    geom_point(x=13.284650+.5, y=52.506761, color="black", size=1, shape=2) +  
    geom_text(x=-9.195222+.5, y=38.743621-.3, label="Lisboa", color="black", size=3) +
    geom_point(x=-9.195222+.5, y=38.743621, color="black", size=1, shape=2) +
    theme(panel.grid.major = element_line(color = gray(.1), linetype = "dotted", size = 0.5),
          panel.background = element_rect(fill = "grey90")) +
    xlab("Longitude") + ylab("Latitude") +
    annotation_scale(location = "br", width_hint = 0.25) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    #xlab("") + ylab("") +
    ggtitle("Population during Romantic period from 1700 to 1750", subtitle = "")+
    labs(color = "Regions", size = "Aggregated Population in 1000")+
    theme(legend.box = "horizontal",
          legend.box.margin = margin(0.2, 0.01, 0.1, 0.01, "cm"),
          legend.margin = margin(1.5, 0.01, 0.01, 0.5, "cm"),
          legend.direction = "vertical",
          legend.spacing = unit(0.01, "cm"),
          legend.text = element_text(size=8),
          legend.title = element_text(size=10))
    # Legend Documentation: https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/


ourPlot
ggsave(file="european_population_period_romantic.png",plot=ourPlot,dpi=300,width=7,height=5)
```

# References

* Mel Moreno, Mathieu Basille. *Drawing beautiful maps programmatically with R, sf and ggplot2 Part 1: Basics*, October 25, 2018, <https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html>.
* Robin Lovelace, Jakub Nowosad, Jannes Muenchow. *Geocomputation with R*, April 24, 2019, <https://geocompr.robinlovelace.net/>, *In particular, check:* Making maps in R, <https://geocompr.robinlovelace.net/adv-map.html>.
* *R Graph Gallery: Maps*, <https://www.r-graph-gallery.com/map/>.
* James Cheshire, *Spatial.ly*, 2016, <http://spatial.ly/r/>.
