---
title: "R Notebook"
output: html_notebook
---

# Graphic devices in R

Plotting works via sending it to different devices. Windows, Linux and Max have different ones to call. 
See ?Devices to find out which ones are available on this computer. 

with command can call a dataset like faithful and direct it to a plot command with x and y arguemtns. 

```{r}
with(faithful, plot(eruptions, waiting))
title(main = "Old Faithful Geyser data")
```

Simple, right?  Now run the command dev.cur(). This will show you the current plotting device, the screen.
```{r}
dev.cur()
```
RStudioGD 
        2 

In case the device is switched to eg. pdf, png, are another other this command will show the device and number. dev.off() will close and revert back to default. 

The second way to create a plot is to send it to a file device. Depending on the type of plot you're making, you explicitly launch a graphics device,
e.g., a pdf file. Type the command pdf(file="myplot.pdf") to launch the file device. This will create the pdf file myplot.pdf in your working directory.
```{r}
pdf(file="myplot.pdf")
with(faithful, plot(eruptions, waiting))
title(main = "Old Faithful Geyser data")
dev.off
```

here are two basic types of file devices, vector and bitmap devices. These use different formats and have different characteristics. Vector formats
| are good for line drawings and plots with solid colors using a modest number of points, while bitmap formats are good for plots with a large number of
| points, natural scenes or web-based plots.

| We'll mention 4 specific vector formats. The first is pdf, which we've just used in our example. This is useful for line-type graphics and papers. It
| resizes well, is usually portable, but it is not efficient if a plot has many objects/points.

| The second is svg which is XML-based, scalable vector graphics. This supports animation and interactivity and is potentially useful for web-based
| plots.
                                                     
| The last two vector formats are win.metafile, a Windows-only metafile format, and postscript (ps), an older format which also resizes well, is usually
| portable, and can be used to create encapsulated postscript files. Unfortunately, Windows systems often don't have a postscript viewer.

| We'll also mention 4 different bitmap formats. The first is png (Portable Network Graphics) which is good for line drawings or images with solid
| colors. It uses lossless compression (like the old GIF format), and most web browsers can read this format natively. In addition, png is good for plots
| with many points, but it does not resize well.

| The device is back to what it was when you started. As you might have guessed, every open graphics device is assigned an integer greater than or equal
| to 2. You can change the active graphics device with dev.set(<integer>) where <integer> is the number associated with the graphics device you want to
| switch to.

| You can also copy a plot from one device to another. This can save you some time but beware! Copying a plot is not an exact operation, so the result
| may not be identical to the original. R provides some functions to help you do this. The function dev.copy copies a plot from one device to another,
| and dev.copy2pdf specifically copies a plot to a PDF file.

 Now run dev.copy with the 2 arguments. The first is png, and the second is file set equal to "geyserplot.png". This will copy your screen plot to a png
| file in your working directory which you can view AFTER you close the device.

```{r}
dev.copy(png, file = "geyserplot.png")
```
png 
  4
  
# Plotting Systems
  
```{r}
# Lattice
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4,1))

```

```{r}
# Base (plot, hist, boxplot, barplot)
with(cars, plot(speed, dist))
text(mean(cars$speed), max(cars$dist), "SWIRL rules!")
```

```{r}
# ggplot2
qplot(displ, hwy, data = mpg)
```

```{r}
#boxplot
boxplot(Ozone~Month, airquality, xlab = "Month", ylab = "Ozone (ppb)", col.axis = "blue", col.lab = "red", main = "Ozone wath ever the f***")
```

```{r}
# histogram
hist(airquality$Ozone)
```

```{r}
# Scatterplot
with(airquality, plot(Wind, Ozone), main = "Ozone...")
```

length with the argument par() now.

[1] 72
```{r}
length(par())
```

| That's correct!
| So there are a boatload (72) of parameters that par() gives you access to. Run the R function names with par() as
| its argument to see what these parameters are.

```{r}
 names(par())
```
 [1] "xlog"      "ylog"      "adj"       "ann"       "ask"       "bg"        "bty"       "cex"       "cex.axis" 
[10] "cex.lab"   "cex.main"  "cex.sub"   "cin"       "col"       "col.axis"  "col.lab"   "col.main"  "col.sub"  
[19] "cra"       "crt"       "csi"       "cxy"       "din"       "err"       "family"    "fg"        "fig"      
[28] "fin"       "font"      "font.axis" "font.lab"  "font.main" "font.sub"  "lab"       "las"       "lend"     
[37] "lheight"   "ljoin"     "lmitre"    "lty"       "lwd"       "mai"       "mar"       "mex"       "mfcol"    
[46] "mfg"       "mfrow"     "mgp"       "mkh"       "new"       "oma"       "omd"       "omi"       "page"     
[55] "pch"       "pin"       "plt"       "ps"        "pty"       "smo"       "srt"       "tck"       "tcl"      
[64] "usr"       "xaxp"      "xaxs"      "xaxt"      "xpd"       "yaxp"      "yaxs"      "yaxt"      "ylbias"   

```{r}
plot(airquality$Wind, airquality$Ozone, type = "n")
points(notmay$Wind,notmay$Ozone,col="red",pch=18)
points(notmay$Wind,notmay$Ozone,col="red",pch=8)
legend("topright", pch=c(17,8), col=c("blue", "red"), legend=c("May", "Other Months"))
abline(v=median(airmiles$Wind), lty=2,lwd=2)
```

```{r}
# plot window for two plots side by side
par(mfrow=c(1,2))
plot(airquality$Wind, airquality$Ozone, main = "Ozone and Wind")
plot(airquality$Ozone, airquality$Solar.R, main = "Ozone and Solar Radiation")
```

```{r}
# setting up the plot window
par(mfrow = c(1,3), mar = c(4,4,2,1), oma = c(0,0,2,0))
plot(airquality$Wind, airquality$Ozone, main = "Ozone and Wind")
plot(airquality$Solar.R, airquality$Ozone, main = "Ozone and Solar Radiation")
plot(airquality$Temp, airquality$Ozone, main = "Ozone and Temperature")
# main title
mtext("Ozone and Weather in New York City", outer = TRUE)
```

# Lattice Plotting System

Plotting systems: scatterplot xyplot(), box-and-whiskers plot bwplot(), boxplots, histogram and other like stripplot, dotplot, splom and levelplot. 

| Here's an example of typical lattice plot call, xyplot(y ~ x | f * g, data). The f and g represent the optional conditioning
| variables. The * represents interaction between them. Remember when we said that lattice is good for plotting multivariate data?
| That's where these conditioning variables come into play.

```{r}
# Example of a lattice basic plot 
xyplot(Ozone~Wind, data = airquality, pch = 8, col = "red", main = "Big Apple Data")
```

```{r}
# latticle plot with a layout that devides it into months
# Since Month is a named column of the airquality dataframe we had to tell R to treat it as a factor.
xyplot(Ozone~Wind | as.factor(Month), data = airquality, layout = c(5,1))
#xyplot(Ozone~Wind | Month, layout = c(5,1), data = airquality) example less informative ?
```

```{r}
#simple plot stored in p
p <- xyplot(Ozone~Wind, data = airquality)
# returns all function of the plot
names(p)
# returns the formular double square brackets are important and quotes
p[["formula"]]
```

```{r}
# lattice plot
xyplot(price~carat | color*cut, data = diamonds, strip = FALSE, pch = 20, xlab = myxlab, ylab = myylab, main = mymain)
```

# Working with Colors
```{r}
# package grDevices
library(grDevices)
# function colors() to contains all available 657 colors
colors() to call all colors. 
# preview of 10 colors
sample(colors(), 10)
#The first, colorRamp, takes a palette of colors (the arguments) and returns a function that takes values between 0 and 1 as
# arguments. The 0 and 1 correspond to the extremes of the color palette. Arguments between 0 and 1 return blends of these extremes.
colorRamp()
# assigning some colors
pal <- colorRamp(c("red", "blue"))
#calling pal with 0
pal(0)
# You see the vector (0,0,255) which represents the highest intensity of blue. What vector do you think the call pal(.5) will return?
# (127.5,0,127.5) since it can only take 0 or 1 the call to .5 return the half of the color

# The argument you pass to the returned function specifies the number of colors you want returned. Each element of the returned vector
# is a 24 bit number, represented as 6 hexadecimal characters, which range from 0 to F. This set of 6 hex characters represents the
# intensities of red, green, and blue, 2 characters for each color.
p1 <- colorRampPalette(c("red", "blue"))
p1(2)
p2 <- colorRampPalette(c("red", "yellow"))
# function to check out
color.scale()
# showMe contains the color.scale function 
showMe(p1(20))
#We mentioned before that colorRamp (and colorRampPalette) could return a 3 or 4 long vector of colors. We saw 3-long vectors
# returned indicating red, green, and blue intensities. What would the 4th entry be?
p3 <- colorRampPalette(c("blue", "green"), alpha = .5)
# simple scattlerplot with assigned rgb colors
plot(x, y, pch = 19, col = rgb(0,.5,.5))
# 4th rgb argument is call to alpha
plot(x, y, pch = 19, col = rgb(0,.5,.5,.3))
```

```{r}
library(RColorBrewer)
#Our last topic for this lesson is the RColorBrewer Package, available on CRAN, that contains interesting and useful color palettes,
# of which there are 3 types, sequential, divergent, and qualitative. Which one you would choose to use depends on your data.

cols <- brewer.pal(3, "BuGn")
pal <- colorRampPalette(cols)
showMe(pal(20))
# provedes a plot with density
image(volcano, col = pal(20))
# different color package
image(volcano, col = p1(20))

# alpha of the rgb() function retruns density
```

