---
title: "R Notebook"
output: html_notebook
---

# Course 8 - ggplot2 with qplot

| The ggplot2 package "is composed of a set of independent components that can be composed in many different ways. ... you can create
| new graphics that are precisely tailored for your problem." These components include aesthetics which are attributes such as colour,
| shape, and size, and geometric objects or geoms such as points, lines, and bars.

```{r}
qplot(displ, hwy, data = mpg, color = drv, geom = c("point", "smooth"))
```

```{r}
qplot(drv, hwy, data = mpg, geom = "boxplot", color= manufacturer)
```

```{r}
qplot(hwy, data = mpg, fill = drv)
```

| It's cool that qplot can do this so easily, but some people may find this multi-color histogram hard to interpret. Instead of using
| colors to distinguish between the drive factors let's use facets or panels. (That's what lattice called them.) This just means we'll
| split the data into 3 subsets (according to drive) and make 3 smaller individual plots of each subset in one plot (and with one call
| to qplot).

| We'll do two plots, a scatterplot and then a histogram, each with 3 facets. For the scatterplot, call qplot with 4 arguments. The
| first two are displ and hwy and the third is the argument data set equal to mpg. The fourth is the argument facets which will be set
| equal to the expression . ~ drv which is ggplot2's shorthand for number of rows (to the left of the ~) and number of columns (to the
| right of the ~). Here the . indicates a single row and drv implies 3, since there are 3 distinct drive factors. Try this now.

```{r}
qplot(displ, hwy, data = mpg, facets = . ~ drv)
```

```{r}
qplot(hwy, data = mpg, facets = drv ~ ., binwidth = 2)
```

# Course 9 ggplot2 with ggplot

| Obviously, there's a DATA FRAME which contains the data you're trying to plot. Then the AESTHETIC MAPPINGS determine how data are
| mapped to color, size, etc. The GEOMS (geometric objects) are what you see in the plot (points, lines, shapes) and FACETS are the
| panels used in conditional plots. You've used these or seen them used in the first ggplot2 (qplot) lesson.

| There are 3 more. STATS are statistical transformations such as binning, quantiles, and smoothing which ggplot2 applies to the data.
| SCALES show what coding an aesthetic map uses (for example, male = red, female = blue). Finally, the plots are depicted on a
| COORDINATE SYSTEM. When you use qplot these were taken care of for you.

```{r}
g <- ggplot(mpg, aes(displ, hwy))
g+geom_point()+geom_smooth(method="lm")
```

```{r}
g+geom_point()+geom_smooth(method="lm")+facet_grid(.~drv)
```

```{r}
g+geom_point()+geom_smooth(method="lm")+facet_grid(.~drv)+ggtitle("Swirl Rules!")
```

# Aesthetics

```{r}
g+geom_point(color = "pink", size = 4, alpha = 1/2)
```

| Now we'll modify the aesthetics so that color indicates which drv type each point represents. Again, use g and add to it a call to
| the function geom_point with 3 arguments. The first is size set equal to 4, the second is alpha equal to 1/2. The third is a call to
| the function aes with the argument color set equal to drv. Note that you MUST use the function aes since the color of the points is
| data dependent and not a constant as it was in the previous example.

```{r}
g+geom_point(size = 4, alpha = 1/2, aes(color = drv))
```

```{r}
g+geom_point(aes(color = drv))+labs(title = "Swirl Rules!")+labs(x = "Displacement", y = "Hwy Mileage")
```

```{r}
g+geom_point(aes(color=drv), size=2, alpha = 1/2)+geom_smooth(size = 4, linetype = 3, method = "lm", se = FALSE)
```
 What did these arguments do? The method specified a linear regression (note the negative slope indicating that the bigger the
| displacement the lower the gas mileage), the linetype specified that it should be dashed (not continuous), the size made the dashes
| big, and the se flag told ggplot to turn off the gray shadows indicating standard errors (confidence intervals).

```{r}
g <- ggplot(testdat, aes(x = myx, y = myy))
g+geom_line()+coord_cartesian(ylim= c(-3,3))
```

```{r}
g <- ggplot(mpg, aes(y = displ, y = hwy, color = factor(year)))
g+geom_point()+facet_grid(drv~cyl, margins = TRUE)
```

| A 4 by 5 plot, huh? The margins argument tells ggplot to display the marginal totals over each row and column, so instead of seeing
| 3 rows (the number of drv factors) and 4 columns (the number of cyl factors) we see a 4 by 5 display. Note that the panel in
| position (4,5) is a tiny version of the scatterplot of the entire dataset.

```{r}
g+geom_point()+facet_grid(drv~cyl, margins = TRUE)+geom_smooth(method = "lm", se = FALSE, size = 2, color = "black")
```

```{r}
g+geom_point()+facet_grid(drv~cyl, margins = TRUE)+geom_smooth(method = "lm", se = FALSE, size = 2, color = "black")+labs(x = "Displacement", y= "Highway Mileage", title = "Swirl Rules!"
```

