---
title: "R Notebook"
output: html_notebook
---



```{r}
library(haven)
bagdad_london_finalRestat <- read_dta("~/R_course/lession4/Bosker_Data/bagdad-london--finalRestat.dta")

#What is the chronological extent of this data?
min(bagdad_london_finalRestat$year)
max(bagdad_london_finalRestat$year)
summary(bagdad_london_finalRestat$year)
```

```{r}
#What periods can it be divided into? How can we do that?
Medieval.bagdad.sub <- subset(bagdad_london_finalRestat,  bagdad_london_finalRestat$year <= 1500)
Medieval.bagdad.sub

EarlyModern.bagdad.sub <- subset(bagdad_london_finalRestat,  bagdad_london_finalRestat$year > 1500)
EarlyModern.bagdad.sub

```

# Can you generate a cumulative graph of population over time, divided into these periods? (Hint: there should be one line for one period and another for another, etc.)

```{r}
Medieval.bagdad.sub <- subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year <= 1500)

EarlyModern.bagdad.sub <- subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year > 1500)

data.frame.Medieval <- data.frame("year" = Medieval.bagdad.sub$year, "country" = Medieval.bagdad.sub$country, "pop" = Medieval.bagdad.sub$total_pop_country)

data.frame.EarlyModern <- data.frame("year" = EarlyModern.bagdad.sub$year, "country" = EarlyModern.bagdad.sub$country, "pop" = EarlyModern.bagdad.sub$total_pop_country)

uniqueMedieval = unique(data.frame.Medieval[c("year", "country", "pop")])
aggMedieval = aggregate(uniqueMedieval$pop, by=list(Category=uniqueMedieval$year), FUN=sum)

uniqueEarlyModern = unique(data.frame.EarlyModern[c("year", "country", "pop")])
aggEarlyModern = aggregate(uniqueEarlyModern$pop, by=list(Category=uniqueEarlyModern$year), FUN=sum)

plot(x = aggMedieval$Category, y = aggMedieval$x, type = "o", col="red", ylab = "Medieval Population", xlab = "Time", main = "Medival Population over Time")
par(new = TRUE)

legend("top", legend=c("Medival"),col=c("red"), lty = 1:2, cex=0.8)

```

```{r}
plot(x = aggEarlyModern$Category, y = aggEarlyModern$x, type = "o", col="green", xlab = "Early Modern Population", ylab = "Time", lty = 2, main = "Early Modern population over Time")
par(new = TRUE)
legend("top", legend=c("Early Modern"),col=c("green"), lty = 1:2, cex=0.8)
```


# North Africa and Europe?
Can you construct comparative graphs of population in North Africa and Europe (similar to what you did with the Morris dataset). Here you will need to sum up population!

```{r}
#North Africa and Europe?
unique_countries <- unique(bagdad_london_finalRestat$country)

Europevector <- c("Austria", "Ireland", "Belgium", "Czech rep.", "Denmark", "Finland", "France", "Germany", "Hungary", "Italy", "Netherlands", "Norway", "Poland", "Portugal", "Slovakia", "Spain", "Sweden", "UK", "Switzerland", "Greece", "Turkey", "Yugoslavia", "Albania", "Bulgaria", "Rumenia", "Luxemburg", "Malta")

NorthAfricavector <- c("Morocco", "Tunesia", "Lybia", "Egypt", "Syria")

Europe.bagdad <- bagdad_london_finalRestat[bagdad_london_finalRestat$country %in% Europevector, ]

NorthAfrica.bagdad <- bagdad_london_finalRestat[bagdad_london_finalRestat$country %in% NorthAfricavector, ]
```

```{r}
data.frame.europe <- data.frame("year" = Europe.bagdad$year, "country" = Europe.bagdad$country, "pop" = Europe.bagdad$total_pop_country)

data.frame.Africa <- data.frame("year" = NorthAfrica.bagdad$year, "country" = NorthAfrica.bagdad$country, "pop" = NorthAfrica.bagdad$total_pop_country)

uniqueAfrica = unique(data.frame.Africa[c("year", "country", "pop")])
aggAfrica = aggregate(uniqueAfrica$pop, by=list(Category=uniqueAfrica$year), FUN=sum)

data.frame.Europe <- data.frame("year" = Europe.bagdad$year, "country" = Europe.bagdad$country, "pop" = Europe.bagdad$total_pop_country)

uniqueEurope = unique(data.frame.Europe[c("year", "country", "pop")])
aggEurope = aggregate(uniqueEurope$pop, by=list(Category=uniqueEurope$year), FUN=sum)

plot(x = aggAfrica$Category, y = aggAfrica$x, type = "l", col="red", ylab = "Population per country", xlab = "Time", main = "Comparative graph of Europe's and North Africa's population")
par(new = TRUE)

plot(x = aggEurope$Category, y = aggEurope$x, type = "l", col="green", axes = FALSE, xlab = " ", ylab = " ", lty = 2)

legend("top", legend=c("North Africa", "Europe"),col=c("red", "green"), lty = 1:2, cex=0.8)

```

```{r}
habsburg <- bagdad_london_finalRestat[bagdad_london_finalRestat$year>=1500, c('city', 'country', 'year')]

exclude_countries <- c("Belgium", "Denmark", "Finland", "France", "Ireland", "Netherlands", "Norway", "Poland", "Portugal", "Spain", "Sweden", "UK", "Switzerland", "Greece", "Turkey", "Lebanon", "Israel", "Syria", "Iraq", "Egypt", "Lybia", "Tunesia", "Algeria", "Morocco", "Yemen", "Saudi Arabia","Oman", "Cyprus", "Luxemburg", "Malta", "Bulgaria")

habsburg <- habsburg[!(habsburg$country %in% exclude_countries), ]


library("dplyr")
habsburg %>% count(habsburg$year)

```


Roxana's fugly "I don't know so I am just tipping" solution to the largest cities in Islamdom per period
```{r}
#Largest cities of Islamdom for each reported period
head(bagdad_london_finalRestat)
acht<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 800)
neun<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 900)
t<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1000)
teins<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1100)
tzwei<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1200)
tdrei<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1300)
tvier<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1400)
tfunf<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1500)
tsechs<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1600)
tsieben<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1700)
tacht<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1800)
#brauch ich auch nicht: Muslimcit<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$muslim == 1)
achtmuslcit<-subset(acht, acht$muslim==1)
max(achtmuslcit$citypop_le10)
neunmuslcit<-subset(neun, neun$muslim==1)
max(neunmuslcit$citypop_le10)
tmuslcit<-subset(t, t$muslim==1)
max(tmuslcit$citypop_le10)
teinsmuslcit<-subset(teins, teins$muslim==1)
max(teinsmuslcit$citypop_le10)
tzweimuslcit<-subset(tzwei, tzwei$muslim==1)
max(tzweimuslcit$citypop_le10)
tdreimuslcit<-subset(tdrei, tdrei$muslim==1)
max(tdreimuslcit$citypop_le10)
tviermuslcit<-subset(tvier, tvier$muslim==1)
max(tviermuslcit$citypop_le10)
tfunfmuslcit<-subset(tfunf, tfunf$muslim==1)
max(tfunfmuslcit$citypop_le10)
tsechsmuslcit<-subset(tsechs, tsechs$muslim==1)
max(tsechsmuslcit$citypop_le10)
tsiebenmuslcit<-subset(tsieben, tsieben$muslim==1)
max(tsiebenmuslcit$citypop_le10)
tachtmuslcit<-subset(tacht, tacht$muslim==1)
max(tachtmuslcit$citypop_le10)

library(dplyr)
cityacht<-filter(achtmuslcit, achtmuslcit$citypop_le10>=350)
cityacht
cityneun<-filter(neunmuslcit, neunmuslcit$citypop_le10>=450)
cityneun
citytausend<-filter(tmuslcit, tmuslcit$citypop_le10>=300)
citytausend
citytausendein<-filter(teinsmuslcit, teinsmuslcit$citypop_le10>=250)
citytausendein
citytausendzwei<-filter(tzweimuslcit, tzweimuslcit$citypop_le10>=200)
citytausendzwei
citytausenddrei<-filter(tdreimuslcit, tdreimuslcit$citypop_le10>=220)
citytausenddrei
citytausendvier<-filter(tviermuslcit, tviermuslcit$citypop_le10>=250)
citytausendfunf<-filter(tfunfmuslcit, tfunfmuslcit$citypop_le10>=280)
citytausendsechs<-filter(tsechsmuslcit, tsechsmuslcit$citypop_le10>=700)
citytausendsieben<-filter(tsiebenmuslcit, tsiebenmuslcit$citypop_le10>=700)
citytausendacht<-filter(tachtmuslcit, tachtmuslcit$citypop_le10>=500)
citytausendacht

```


```{r}
acht
achtmuslcit
```


# Subsetting data

## Introducing North Africa and Europe
Filter for Countries with unique save it as a vector in a variable and create csv with country plus continents. 
Than merge this table with the original datafile in R (variable). R can merge it automatically. (find code)

## Introducing Austrian-Hungarian Empire

Filter countries and assign oberservation "Austrian..." in a variable called Dynastie.
Include countries of the empire and reduce the dataset to this countries.
Exclude years before the empire existed.
Exclude cities that were never part of the empire.
Define cities that were part of the empire from to. 
